The imprint of global magma oceans on exoplanet demographics

### 2023-03-20 src/tex/ms.tex
introduce caveats, update ToDos

% Define document class
\documentclass[twocolumn]{aastex631}
%\documentclass[modern]{aastex631}

@@

% Main body
\section{Introduction}
...just dropping material and references for now...
\todo[inline]{papers: \citep[][introduce runaway GH]{} \citep[][statistical detection of GH transition]{Turbet2019}, \citep[][(initial) water content of terrestrial planets]{Lichtenberg2019} \citep[][water loss, O2 buildup around M dwarfs]{Luger2015}, \citep[][Solubility of H in magma oceans]{Hirschmann2012,SchlichtingShahar_inreview,} \citet{Hamano2013,Hamano2015,Barth2021,Downey2022}.}
\todo[inline]{``Points for the intro:
\bioverse\\
Transit survey to measure mass/radius of planets\\
Does the HZ have an inner edge?\\
Run-away GH\\
Magma oceans\\
Mention missions that are NOT statistical and therefore not considered here: CHEOPS, ARIEL, Roman. Other missions/concepts do not directly measure radii and not considered here: Habitable Worlds Observatory, LIFE.\\
~\\
In contrast, PLATO and \nautilus\ are \\
Possibility of follow-up to measureme masses
''}

@@

With this updated distribution function, planets are generated as in the original \bioverse\ paper, but with an additional stellar mass dependence (for a star of a given mass, the relevant model parameters from the bin containing that mass are used). In summary, stars are assigned a number of planets in accordance with $F_0$, with planet parameters randomly drawn in proportion to the occurrence distribution function evaluated over a grid of radii and orbital periods.
\todo[inline]{GALEN: The most relevant assumption of using our model is that, at the relevant fluxes (I think), we predict that super-Earths are pretty common. I think the "drop" in our model where super-Earths become less common happens at later fluxes than what is relevant in this paper... Let's double check this, and add a note in the first paragraph about it?.}

@@

The measured radii $R_\mathrm{P, i}$ cannot be directly used for the hypothesis tests as they include intrinsic scatter that is not caused by measurement errors.
$H_{\mathrm{rgh}}$ and $H_0$ should thus be tested against a statistical estimator that represents the population mean.
To avoid binning and the artificial patterns it may introduce, we chose to test our hypotheses against a simple moving average $SMA$ along the instellation axis with a window of size \windowsize\ centered around each measurement.\footnote{See Appendix~\ref{app:binnedstats} for a robustness test using a different estimator.}
We further computed the uncertainty of this moving average by propagating the individual measurement errors and applying a rolling standard error of the mean. \todo[inline]{specify error propagation}

@@

\subsection{Mission design trades}\label{sec:mission-design-trades}
\todo[inline]{not sure if subsubsections needed}
To explore the impact of mission trade-offs on the detectability of the runaway greenhouse transition, we simulated different survey designs and strategies and measured their capability to recover the trend and constrain its parameters.
We assessed this capability based on two determinants: the likelihood that the mission is able to detect the injected trend, and the precision with which it can constrain the parameters of that trend.

@@

\subsection{Constraining planetary habitability}\label{sec:habitability}
\todo[inline]{Magma oceans influence the amount of water available at the surface and atmosphere, which is 1. commonly used as environmental marker to assess habitability, and 2. influences the planet's climate (greenhouse effect).}
A potential for liquid water on the surface of a planet is commonly used as an environmental marker to assess its habitability~\citep{Huang1959,Hart1978,Kasting1993,Kaltenegger2011,Kopparapu2013}.
The runaway greenhouse transition represents an upper bound on received irradiation for this condition.
Its detection would thus not only empirically confirm the habitable zone concept but also help to locate it in the observationally available planetary parameter space.
In Sect.~\ref{sec:res_constrain-S_thresh}, we show that the threshold instellation at which the runaway greenhouse transition occurs can be reasonably constrained without imposing overly optimistic conditions on the underlying planet population, instrumentation, or survey strategy.
A mission like \plato\ is well equipped to perform this measurement; the chances for success are directly proportional to the proportion of characterized planets around M dwarfs and to the number of planets for which masses can be determined.

@@

\subsection{Impact of assumptions on our findings}
\todo[inline]{Caveats\\
context: \citet{Turbet2019} Sect. 3, \citet{Bower2019}.\\
''present bracketing cases; or to discuss why the discountinuity in the results would not be affected by changes that led to continuous offsets in the results. Or both.''\\
- not yet explored: how sensitive are we on the assumed planet population? $\exists$ danger of false positive detection on occurrence rate ``ridges'' like the Neptune desert? Currently we have a relatively smooth occurrence rate density in $R_P - S$ space based on the power-law fits in \citet{Bergsten2022}. We could, with relatively moderate effort, run a sanity check on an alternative sample (e.g., the more structured Bern model populations, or the population in the Genesis database)}

>>>>

% Define document class
\documentclass[twocolumn,twocolappendix]{aastex631}
%\documentclass[modern]{aastex631}

@@

% Main body
\section{Introduction}
\todo[inline]{papers: \citep[][introduce runaway GH]{} \citep[][statistical detection of GH transition]{Turbet2019}, \citep[][(initial) water content of terrestrial planets]{Lichtenberg2019} \citep[][water loss, O2 buildup around M dwarfs]{Luger2015}, \citep[][Solubility of H in magma oceans]{Hirschmann2012,SchlichtingShahar_inreview,} \citet{Hamano2013,Hamano2015,Barth2021,Downey2022}.}
\todo[inline]{``Points for the intro:
\bioverse\\
Transit survey to measure mass/radius of planets\\
Does the HZ have an inner edge?\\
Run-away GH\\
Magma oceans\\
Mention missions that are NOT statistical and therefore not considered here: CHEOPS, ARIEL, Roman. Other missions/concepts do not directly measure radii and not considered here: Habitable Worlds Observatory, LIFE.\\
~\\
In contrast, PLATO and \nautilus\ are \\
Possibility of follow-up to measureme masses
''}

@@

With this updated distribution function, planets are generated as in the original \bioverse\ paper, but with an additional stellar mass dependence (for a star of a given mass, the relevant model parameters from the bin containing that mass are used). In summary, stars are assigned a number of planets in accordance with $F_0$, with planet parameters randomly drawn in proportion to the occurrence distribution function evaluated over a grid of radii and orbital periods.

@@

The measured radii $R_\mathrm{P, i}$ cannot be directly used for the hypothesis tests as they include intrinsic scatter that is not caused by measurement errors.
$H_{\mathrm{rgh}}$ and $H_0$ should thus be tested against a statistical estimator that represents the population mean.
To avoid binning and the artificial patterns it may introduce, we chose to test our hypotheses against a simple moving average $SMA$ along the instellation axis with a window of size \windowsize\ centered around each measurement.\footnote{See Appendix~\ref{app:binnedstats} for a robustness test using a different estimator.}
We further computed the uncertainty of this moving average by propagating the individual measurement errors and applying a rolling standard error of the mean. %\todo[inline]{specify error propagation}

@@

\subsection{Mission design trades}\label{sec:mission-design-trades}
To explore the impact of mission trade-offs on the detectability of the runaway greenhouse transition, we simulated different survey designs and strategies and measured their capability to recover the trend and constrain its parameters.
We assessed this capability based on two determinants: the likelihood that the mission is able to detect the injected trend, and the precision with which it can constrain the parameters of that trend.

@@

\subsection{Constraining planetary habitability}\label{sec:habitability}
%\todo[inline]{Magma oceans influence the amount of water available at the surface and atmosphere, which is 1. commonly used as environmental marker to assess habitability, and 2. influences the planet's climate (greenhouse effect).}
A potential for liquid water on the surface of a planet is commonly used as an environmental marker to assess its habitability~\citep{Huang1959,Hart1978,Kasting1993,Kaltenegger2011,Kopparapu2013}.
The runaway greenhouse transition represents an upper bound on received irradiation for this condition.
Its detection would thus not only empirically confirm the habitable zone concept but also help to locate it in the observationally available planetary parameter space.
In Sect.~\ref{sec:res_constrain-S_thresh}, we show that the threshold instellation at which the runaway greenhouse transition occurs can be reasonably constrained without imposing overly optimistic conditions on the underlying planet population, instrumentation, or survey strategy.
A mission like \plato\ is well equipped to perform this measurement; the chances for success are directly proportional to the proportion of characterized planets around M dwarfs and to the number of planets for which masses can be determined.

@@

\subsection{Impact of assumptions on our findings}
\todo[inline]{Caveats\\
context: \citet{Turbet2019} Sect. 3, \citet{Bower2019}.\\
''present bracketing cases; or to discuss why the discountinuity in the results would not be affected by changes that led to continuous offsets in the results. Or both.''\\
- not yet explored: how sensitive are we on the assumed planet population? $\exists$ danger of false positive detection on occurrence rate ``ridges'' like the Neptune desert? Currently we have a relatively smooth occurrence rate density in $R_P - S$ space based on the power-law fits in \citet{Bergsten2022}. We could, with relatively moderate effort, run a sanity check on an alternative sample (e.g., the more structured Bern model populations, or the population in the Genesis database)}
The prospects for detecting the runaway greenhouse transition depends on astro- and geophysical factors, as well as on the specific instrumentation and survey strategy of a particular mission.
We can only approximate all this here, and we therefore seek to assess the sensitivity of our results on the assumptions we have made.
